### 2022-04-08 outputs/paper/paper.Rmd
add abstract

---
title: "Education and region effects on Philippines teenage fertility in 1998"
author: 
  - Zhongyu Huang
thanks: "Code and data are available at: https://github.com/zhongyuhuang/DHS_Analysis-.git"
date: "April 6, 2022"
abstract: "Teenage fertilities induce health risks and life impacts on both young mothers and their children. Philippines, one of the countries that have the highest adolescent birth rates, the teenage pregnancy toll is still rising here. This paper uses data collected by the National demographic and health survey in 1998 to investigate how background characteristics affect the teenage fertility rates in the Philippines. Evidence has shown that most teenage mothers received less education and lived in rural areas."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
# setup workspace
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2)
library(gridExtra)
```

```{r load, include=FALSE}
# load data
f <- file.choose()
dhs1998 <- read.csv(f)
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Percentage of teenage pregnancy and motherhood by residence", echo = FALSE, fig.height=9, fig.width=5}
#subset the dataset by rows containing residence
Residence<- dhs1998[8:9,]

# plot Percentage of who are mothers
p1 <- ggplot(Residence, aes(BG_inf, PT_Mom)) + xlab("Residence") +ylab("Percentage of who are mothers")+geom_col()

# plot Percentage of who have begun child-bearing
p2 <- ggplot(Residence, aes(BG_inf, PT_bear)) + xlab("Residence") +ylab("Percentage of who have begun child-bearing")+geom_col()

# plot Percentage of who are pregnant with first child
p3 <- ggplot(Residence, aes(BG_inf, PT_Pgt)) + xlab("Residence") +ylab("Percentage of who are pregnant with first child")+geom_col()

# arrange plots
grid.arrange(p1,p2,p3)

```

Our data is of penguins (Figure \@ref(fig:bills2)).

```{r bills2, fig.cap="Percentage of teenage pregnancy and motherhood by region", echo = FALSE, fig.height=9, fig.width=5}
#subset the dataset by rows containing age
Region <- dhs1998[11:26,]

# plot Percentage of who are mothers
c1 <- ggplot(Region, aes(BG_inf, PT_Mom)) + xlab("Region") +ylab("Percentage of who are mothers")+ geom_col()+coord_flip()

# plot Percentage of who have begun child-bearing
c2 <- ggplot(Region, aes(BG_inf, PT_bear)) + xlab("Region") +ylab("Percentage of who have begun child-bearing")+ geom_col()+coord_flip()

# plot Percentage of who are pregnant with first child
c3 <- ggplot(Region, aes(BG_inf, PT_Pgt)) + xlab("Region") +ylab("Percentage of who are pregnant with first child")+ geom_col() +coord_flip()

# arrange plots
grid.arrange(c1,c2,c3)
```

```{r bills3, fig.cap="Percentage of teenage pregnancy and motherhood by education level", echo = FALSE, fig.height=9, fig.width=5}
#subset the dataset by rows containing age
Education<- dhs1998[28:31,]

# plot Percentage of who are mothers
p1 <- ggplot(Education, aes(BG_inf, PT_Mom))+ xlab("Education Level") +ylab("Percentage of who are mothers") + geom_col()

# plot Percentage of who have begun child-bearing
p2 <- ggplot(Education, aes(BG_inf, PT_bear))+ xlab("Education Level") +ylab("Percentage of who have begun child-bearing") +geom_col()

# plot Percentage of who are pregnant with first child
p3 <- ggplot(Education, aes(BG_inf, PT_Pgt)) + xlab("Education Level") +ylab("Percentage of who are pregnant with first child") +
  geom_col()

# arrange plots
grid.arrange(p1,p2,p3)

```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References